---
title: "Urban population growth prediction"
output:
  word_document: default
  html_document: default
date: "2024-10-15"
---
*Urban Population Growth Forecasting using ARMA*

1. Data Import and Transformation
```{r}
file.exists("C:\\Users\\mahak\\Downloads\\urban_pop(1986-2019).xls")
data<- readxl::read_excel("C:\\Users\\mahak\\OneDrive\\Documents\\R-Script\\urban_pop(1986-2019).xls")
data
```
```{r}
urban_pop1<- ts(data[,2], start= 1986, end = 2019)
urban_pop1
```

2. Visualization
```{r}
library(fpp2)
autoplot(urban_pop1)+xlab("Time")+ylab("Urban population growth")+ggtitle("Urban Population Growth in India(1986-2019)")
```
3. Checking for Time Trend and Seasonality
Lagplot:
```{r}
gglagplot(urban_pop1)
```

ACF values:
```{r}
acf_values <- acf(urban_pop1, lag.max=9, plot=F)
acf_values$acf
```

Plotting the correlogram:
```{r}
ggAcf(urban_pop1)
```
4. Model Estimation

not that: we are finding the best fit model using non-stationary data
Plotting all the forecasting methods:

```{r}
autoplot(urban_pop1)+
  autolayer(meanf(urban_pop1,h=20), series="Mean",PI=T)+
  autolayer(naive(urban_pop1,h=20), series="Naive",PI=T)+
  autolayer(snaive(urban_pop1,h=20), series="Seasonal naive",PI=T)+
  autolayer(rwf(urban_pop1,h=20,drift = T), series="Drift",PI=T)
```

Plotting the residual graph:

1.*Using naive method:*
```{r}
res<- residuals(naive(urban_pop1))
autoplot(res)+xlab("Year")+ylab("")+ggtitle("Residuals from naive method")
```

Plotting the histogram of residuals:

```{r}
gghistogram(res)+ggtitle("Histogram of Residuals for Urban Population Growth")
```


```{r}
checkresiduals(naive(urban_pop1))
```



2. *Using seasonal naive method:*

```{r}
res1<- residuals(snaive(urban_pop1))
autoplot(res1)+xlab("Year")+ylab("")+ggtitle("Residuals from seasonal-naive method")
```


plotting the residual histogram:

```{r}
gghistogram(res1)+ggtitle("Histogram of residuals")
```


all in one graph:
```{r}
checkresiduals(snaive(urban_pop1))
```


6. Model Evaluation 
Evaluating forecasting accuracy:


```{r}
urban_pop_training_data<- window(urban_pop1, end=2009) #training data
urbanpop1<-  meanf(urban_pop_training_data, h=10)
urbanpop2<- rwf(urban_pop_training_data, h=10)
urbanpop3<- snaive(urban_pop_training_data, h=10)
urbanpop4<- rwf(urban_pop_training_data, drift = T, h=10)

autoplot(urban_pop1)+ #plotting all the data
  autolayer(urbanpop1, series="Mean", PI= F)+
  autolayer(urbanpop2, series="Naive", PI= F)+
  autolayer(urbanpop3, series="Seasonal Naive", PI= F)+
   autolayer(urbanpop4, series="Drift", PI= F)+
  xlab("Year")+ylab("Urban population growth(%)")+
  ggtitle("Forecasts for urban population growth")+
  guides(colour= guide_legend(title="Forecast"))
```


Checking the accuracy:
1.mean forecasting method:
```{r}
urban_pop_test_data<- window(urban_pop1, start=2010) #test data
accuracy(urbanpop1, urban_pop_test_data)

```
2.Naive method
```{r}
accuracy(urbanpop2, urban_pop_test_data)

```

3.Seasonal naive method
```{r}
accuracy(urbanpop3, urban_pop_test_data)

```

4.Drift method
```{r}
accuracy(urbanpop4, urban_pop_test_data)
```


now, converting the non-stationary data into stationary data and then comparing the two models:
Checking for stationarity:

```{r}
cbind("Urban population growth(%)"= urban_pop1, "Monthly log urban population growth"= log(urban_pop1), "Yearly change in log urban popn growth"= diff(log(urban_pop1),12)) %>%
  autoplot(facets=T)+xlab("Year")+ylab("")+ggtitle("Urban population growth rate in different log transformations:")
```


test:
```{r}
library(urca)
urban_pop1 %>% ur.kpss() %>% summary()
```

since the data is non-stationary, we need to finf the best fitted model to convert it into a stationary series.
which model to use?
```{r}
summary((fit<- auto.arima(urban_pop1)))
```

We use auto.arima() to get the best fitted model. Here the best fitted model is ARIMA(0,1,0)
Note that: The RMSE value when the data is converted to a stationary series is less than the previous RMSE value where we directly forecasted the data.

now we will forecast the next 10 years:
Forecasting:
```{r}
fit%>% forecast(h=10)%>% autoplot(include=20)
```


```{r}
checkresiduals(fit)
```




